\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-02-07 paper/main.tex
Results restructuring

Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model. Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously. By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.

\subsection{\name identifies drugs that target a protein}

\todo[inline]{Likely discussion:}
The aim of \name is to predict candidate drugs that target a given
protein; the challenge is to develop a training and evaluation scheme
that does not simply overfit to the inherent biases in training and
testing data.
In general, when performing cross-validation for DTI prediction, the
options are to split over 
\begin{enumerate}
	\item split over drugs,
	\item split over drug--target pairs, or
	\item split over proteins
\end{enumerate}
where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions. They ensure that at least parts of the interactions are not seen during training and evaluate either how well targets are predicted for unseen drugs or unseen proteins. Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model. \\

Assuming a hypothetical, perfectly generalizing model built upon an unbiased dataset, this very model would yield similar performances for all three splitting schemes, not overfitting on the known structures. On the other hand, for a hypothetical entirely overfitting model, trained on a highly biased dataset, this model would show substantial deviations from the original performance over another split. \\
We emphasize, that all real-world models are prone to some sort of overfitting, and unknown, deviant entities in both validation and testing set will likely lead to some sort of performance gap for relevant metrics. However, a large disparity may hint the biases stated above.

Hence, we will perform a cross-validation over proteins for the training and prediction phase of our model, despite predicting per protein is rather counter-intuitive as there only limited drug-targets \citep{Overington2006}, and thus novel drugs are more likely to arise than novel targets. Yet, we aim to find all interacting drugs for existing targets motivating our split choice even further.

\subsection{Experiments}
In this section we will present the results of \name on both STITCH and Yamanishi benchmark dataset, omitting any kind of comparison for now. We trained, validated and finally tested all considered models on STITCH dataset using a 5-fold cross-validation over a protein split, as reasoned and substantiated previously. We only evaluated the best performing model, with respect to \textit{AUROC} and $MicroAUC_p$, on the Yamanishi benchmark dataset. 
We will hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}. As depicted in Figure \ref{tab:STITCHresults}, \textit{OntoPred} is showing significantly better performance on STITCH, while also only \textit{OntoPred} shows significant performance increase when adding \textit{GENConv} graph convolutional neural layers over the PPI graph. However, graph convolutional neural layers such as \textit{GCNConv} and \textit{GENConv}, especially when incorporated into the proposed \textit{ResGraphConv} blocks, add a lot of learnable parameters to the network, leading to more expressive power. This may lead to better performance, while deceiving, whether the respective features actually localize on the PPI graph.

@@

>>>>

We first perform a cross-validation over proteins and validate our
results. Cross-validation over proteins aims to evaluate how the model
performs when tasked to identify drugs that may target a ``novel''
protein, i.e., one not seen during training.
\todo[inline]{Describe/add the CV results here.}

Second, we use the model trained on STITCH and evaluate the model on
the Yamanishi benchmark dataset\todo{This needs another sentence
  stating how you made sure there was no overlap in data.}.

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein
split. We only evaluated the best performing model (with respect to
\textit{AUROC} and $MicroAUC_p$) on the Yamanishi benchmark dataset.

We will hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}. As depicted in Figure \ref{tab:STITCHresults}, \textit{OntoPred} is showing significantly better performance on STITCH, while also only \textit{OntoPred} shows significant performance increase when adding \textit{GENConv} graph convolutional neural layers over the PPI graph. However, graph convolutional neural layers such as \textit{GCNConv} and \textit{GENConv}, especially when incorporated into the proposed \textit{ResGraphConv} blocks, add a lot of learnable parameters to the network, leading to more expressive power. This may lead to better performance, while deceiving, whether the respective features actually localize on the PPI graph.

@@

% despite predicting per protein is
% rather counter-intuitive as there only limited drug-targets
% \citep{Overington2006}, and thus novel drugs are more likely to arise
% than novel targets. Yet, we aim to find all interacting drugs for
% existing targets motivating our split choice even further.

\todo[inline]{Methods:}
Through the very nature of the graph convolutional neural network, we
build the transformed representation for all proteins in every
forwarding step of the model. Note particularly, that the same
convolutional filter and weight matrix are applied to and learned for
all nodes simultaneously. By construction, for a single drug we can
compute and predict all its interactors in a single run of the model,
leading to significantly less computing time.

%\subsection{\name identifies drugs that target a protein}

\todo[inline]{Likely discussion:}
The aim of \name is to predict candidate drugs that target a given
protein; the challenge is to develop a training and evaluation scheme
that does not simply overfit to the inherent biases in training and
testing data.
In general, when performing cross-validation for DTI prediction, the
options are to split over 
\begin{enumerate}
	\item split over drugs,
	\item split over drug--target pairs, or
	\item split over proteins
\end{enumerate}
where the first and third option concern splitting drugs and proteins,
respectively, into train, validation and test sets, and arranging the
corresponding drug-target interactions. They ensure that at least
parts of the interactions are not seen during training and evaluate
either how well targets are predicted for unseen drugs or unseen
proteins. Hereby, different training and prediction schemes lead to
divergent expressiveness of the resulting model.

\todo[inline]{Likely discussion:}
Assuming a hypothetical, perfectly generalizing model built upon an
unbiased dataset, this very model would yield similar performances for
all three splitting schemes, not overfitting on the known
structures. On the other hand, for a hypothetical entirely overfitting
model, trained on a highly biased dataset, this model would show
substantial deviations from the original performance over another
split.

\todo[inline]{Likely discussion:}
We emphasize, that all real-world models are prone to some sort of
overfitting, and unknown, deviant entities in both validation and
testing set will likely lead to some sort of performance gap for
relevant metrics. However, a large disparity may hint the biases
stated above.

\subsection{Experiments}
